\documentclass[prodmode,acmtcps]{acmsmall}

% Metadata Information
\acmVolume{9}
\acmNumber{4}
\acmArticle{39}
\acmYear{2016}
\acmMonth{7}

\usepackage{graphicx}
%\graphicspath{{figs/}}
\usepackage{paralist}
\usepackage{todonotes}  %\newcommand{\todo}[][a]{}
\usepackage{color}
\usepackage{epstopdf}
\usepackage{subfigure}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{cool}  % COntext-Oriented math for LaTeX
\usepackage{mleftright}  % Commands \mleft and \mright
\usepackage{booktabs}
%\usepackage{subfloat}
\usepackage[per-mode=symbol]{siunitx}
\usepackage{url}


\usepackage{hyperref}
\hypersetup{citecolor=black, colorlinks=false}
\usepackage[font=small]{caption}
\usepackage{cite}

% Also, note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
%\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks
\usepackage{comment}
\usepackage[amsmath,thmmarks,thref]{ntheorem}
\usepackage{algorithm,algpseudocode}
\theoremstyle{definition}
\newtheorem{observation}{Observation}
\usepackage{truonglatexdefs}

\usepackage{enumitem, kantlipsum}
\usepackage{tikz}
\usetikzlibrary{arrows}
\tikzset{
  treenode/.style = {align=center, inner sep=0pt, text centered},
  branch_c/.style = {treenode, circle, black, draw=black,
    text width=1.5em, very thick},  
  branch_l/.style = {treenode, circle, blue!50!gray!, draw=blue!50!gray!,
    text width=1.5em, very thick},
  branch_r/.style = {treenode, circle, red!50!gray!, draw=red!50!gray!, 
    text width=1.5em, very thick},
}

%\usepackage{balance}
\def\u{\mathrm{u}}
\def\x{\mathrm{x}}
\def\y{\mathrm{y}}

%\usepackage{etoolbox}
%\makeatletter
%\patchcmd{\maketitle}{\@copyrightspace}{}{}{}
%\makeatother

%\apptocmd{\thebibliography}{\small}{}{}

\begin{document}
% Page heads
\markboth{A. Jain et al.}{Data-driven Model Predictive Control with Regression Trees -- \\
	An Application to Building Energy Management}

% Title portion
\title{Data-driven Model Predictive Control with Regression Trees --\\
	An Application to Building Energy Management}
\author{ACHIN JAIN
\affil{University of Pennsylvania}
FRANCESCO SMARRA
\affil{Universit\`{a} degli Studi dell'Aquila}
MADHUR BEHL
\affil{University of Virginia}
RAHUL MANGHARAM
\affil{University of Pennsylvania}}

% NOTE! Affiliations placed here should be for the institution where the
%       BULK of the research was done. If the author has gone to a new
%       institution, before publication, the (above) affiliation should NOT be changed.
%       The authors 'current' address may be given in the "Author's addresses:" block (below).
%       So for example, Mr. Abdelzaher, the bulk of the research was done at UIUC, and he is
%       currently affiliated with NASA.

\begin{abstract}

\textcolor[rgb]{0.00,0.00,1.00}{Large-scale cyber-physical systems optimization is one of the most important problems nowadays, therefore the use of optimal control strategies is fundamental. Model Predictive Control (MPC) is a control strategy based on the ability to predict system's behavior, that provides optimal solutions in order to optimize a generic system performance while ensuring system's constraints. For this reason MPC represents the best solution to optimize systems' performance. However, the main bottle neck that does not allow the widespread of predictive control algorithms for complex systems is mainly due to their complexity. This is because the resulting cost and time efforts needed to construct system models, as for example in the case of buildings, is prohibitive. To address this issue, a regression trees-based approach is introduced in this paper. More precisely, a data-driven model predictive control based on regression trees is proposed, which allows to perform closed-loop optimal control synthesis for cyber-physical systems. In this paper, such methodology is used to face the Demand-Response problem in buildings. In particular, it is used to optimally trade off peak power reduction against thermal comfort without having to learn white/grey box models of the systems dynamics.}
\end{abstract}


%
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below. 
%
\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10010147.10010257.10010293.10003660</concept_id>
<concept_desc>Computing methodologies~Classification and regression trees</concept_desc>
<concept_significance>500</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Computing methodologies~Classification and regression trees}


%
% End generated code
%

% We no longer use \terms command
%\terms{Design, Algorithms, Performance}

\keywords{Machine learning; Predictive control; Cyber-physical Systems; Demand response; Peak power reduction}

\acmformat{Achin Jain, Francesco Smarra, Madhur Behl and Rahul Mangharam, 2016. Data-driven Model Predictive Control with Regression Trees -- An Application to Building Energy Management.}

% At a minimum you need to supply the author names, year and a title.
% IMPORTANT:
% Full first names whenever they are known, surname last, followed by a period.
% In the case of two authors, 'and' is placed between them.
% In the case of three or more authors, the serial comma is used, that is, all author names
% except the last one but including the penultimate author's name are followed by a comma,
% and then 'and' is placed before the final author's name.
% If only first and middle initials are known, then each initial
% is followed by a period and they are separated by a space.
% The remaining information (journal title, volume, article number, date, etc.) is 'auto-generated'.

\begin{bottomstuff}
Author's addresses: {A. Jain {and} R. Mangharam, Department of Electrical and Systems Engineering,
University of Pennsylvania; 
F. Smarra, Department of Information Engineering, Computer Science and Mathematics, Universit\`{a} degli Studi dell'Aquila;
M. Behl, Department of Computer Science, University of Virginia.
This work was partially supported by the Italian Government under Cipe resolution n.135 (Dec. 21, 2012), project \emph{INnovating City Planning through Information and Communication Technologies} (INCIPICT).}
\end{bottomstuff}

\maketitle

\section{Introduction}
\label{sec:intro}
\input{introduction.tex}

\section{Problem Definition}
\label{sec:problem}
\input{problem.tex}

\section{Control with Regression Trees: Background}
\label{sec:drtree}
\input{DRsynthesis.tex}

\section{DPCRT: Data-driven Predictive Control using Regression Trees}
\label{sec:decision_tree}
\input{DecisionTrees.tex}

\section{Case Study: Peak Power Reduction}
\label{sec:case}
\input{case.tex}

\section{Conclusion}
\label{sec:discussion}
\input{conclusion.tex}


%\let\secfnt\undefined
%\newfont{\secfnt}{ptmb8t at 10pt}

\bibliographystyle{unsrt}
\bibliography{references,ref_buildsys,newrefs}

\end{document}


